Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-03-15 paper/tex/analysis.tex
consistent messaging around figure 1
fixing a bug I introduced

\subsection{Metric of individual photo-\mathinhead{$z$}{} posteriors}
\label{sec:CDE_loss}
%(Alex Malz, Ann Lee)

@@

\subsection{Metrics of an ensemble of photo-\mathinhead{$z$}{} posteriors}
\label{sec:qualmet}
%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
%(Alex Malz, Peter Freeman, Sam Schmidt)
%(John Soo, Rongpu Zhou)

@@

Perhaps the most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, a quantity that enters some cosmological calculations and the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
This is the exact estimator, \trainz, that we have described in Section~\ref{sec:method:trainz}, and which will serve as an experimental control.

>>>>

\subsection{Metric of individual \pzpdf s}
\label{sec:CDE_loss}
%(Alex Malz, Ann Lee)

@@

\subsection{Metrics of an ensemble of \pzpdf s}
\label{sec:qualmet}
%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
%(Alex Malz, Peter Freeman, Sam Schmidt)
%(John Soo, Rongpu Zhou)

@@

Perhaps the most popular application of \pzpdf s is the estimation of the overall redshift distribution $N(z)$, a quantity that enters some cosmological calculations and the true value of which is known for the DC1 data set and will be denoted as $\tilde{N}(z)$.
In terms of the prior information provided to each method, the true redshift distribution satisfies the tautology $\tilde{N}(z) = p(z \vert I_{D})$ due to our experimental set-up; because the DC1 training and template sets are representative and complete, $I_{D}$ represents a prior that is also equal to the truth.
In this ideal case of complete and representative prior information, the method that would give the best approximation to $\hat{N}(z)$ would be one that neglects all the information contained in the photometry $\{d_{i}\}_{N_{tot}}$ and gives every galaxy the same \pzpdf\ $\hat{p}_{i}(z) = \tilde{N}(z)$ for all $i$; the inclusion of any information from the photometry would only introduce noise to the optimal result of returning the prior.
This is the exact estimator, \trainz, that we have described in Section~\ref{sec:trainz}, and which will serve as an experimental control.

### 2019-03-15 paper/tex/discussion.tex
fixing a bug I introduced

The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
As mentioned in Section~\ref{sec:stackedmetrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
The effects of tomographic bin selection will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.

>>>>

The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
As mentioned in Section~\ref{sec:metrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
The effects of tomographic bin selection will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.

### 2019-03-15 paper/tex/results.tex
fixing a bug I introduced

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
\caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of $0$ and $1$, and asymmetry indicates systematic bias, a form of model misspecification.}
\label{fig:pitqq}
\end{figure*}

@@

Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with a dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
\textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.
The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.
A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

@@

\subsection{Metrics of the stacked estimator of the redshift distribution}\label{sec:stackedmetrics}

@@

%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes}
%\textcolor{cyan}{Ann: }

@@

\subsection{Interpretation of metrics}\label{sec:caution}
%Caution on metrics and photo-$z$ codes}\label{sec:caution}
%(Alex Malz, Sam Schmidt)

@@

The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:method:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
%\textit{Explain the connection between \trainz\ and these metrics.}

>>>>

\begin{figure*}
\centering
\includegraphics[width=0.74\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
\caption{The QQ plot (red) and PIT histogram (blue) of the \pzpdf\ codes (panels) along with the ideal QQ (black dashed diagonal) and ideal PIT (gray horizontal) curves, as well as a difference plot for the QQ difference from the ideal diagonal (lower inset).
The twelve codes exhibit varying degrees of four deviations from perfection: an overabundance of PIT values at the centre of the distribution indicate a catalogue of overly broad \pzpdf s, an excess of PIT values at the extrema indicates a catalogue of overly narrow \pzpdf s, catastrophic outliers manifest as overabundances at PIT values of 0 and 1, and asymmetry indicates systematic bias, a form of model misspecification.}
\label{fig:pitqq}
\end{figure*}

@@

Examination of the PIT histograms and QQ plots smethod:hows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with a dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.
\textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.
In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.
The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.
A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

@@

\subsection{Metrics of the stacked estimator of the redshift distribution}
\label{sec:stackedmetrics}

@@

%\red{discussion of why NN and FZBoost are doing well, how we can try to improve the other codes}

@@

\subsection{Interpretation of metrics}
\label{sec:caution}
%Caution on metrics and photo-$z$ codes}\label{sec:caution}
%(Alex Malz, Sam Schmidt)

@@

The \trainz\ estimator, which assigns every galaxy a $p(z)$ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a ``null test'' to demonstrate this point via \textit{reductio ad absurdum}.
\trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
%\textit{Explain the connection between \trainz\ and these metrics.}
